---
title: "Metapopulation Importation Calculator for Seasonal Influenza"
author: "Jose Herrera-Diestra"
date: "`r format(Sys.Date())`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Overview

This R Markdown provides a **complete, reproducible** importation calculator for seasonal influenza. It estimates:

1. Expected **infectious introductions ("seeds")** from international travelers (including those incubating who become infectious during their stay).
2. **Expected total cases** generated by those seeds over a finite horizon using a branching-process approximation, with **different** $R_t$ for community vs. micro-settings.
3. **Probability of ≥1 self-sustaining chain** (when $R_t > 1$) using a Negative Binomial (overdispersed) offspring model.
4. (Optional) **Monte Carlo** summaries of per-seed cluster sizes.

> **Use case**: Large numbers of travelers arrive to the U.S. in June–July from the Southern Cone (Argentina, Chile, southern Brazil) when their influenza season peaks.

# Assumptions

- **Incubation**: Lognormal (median ~1–2 days for influenza).
- **Generation time**: ~2–3 days.
- **Offspring**: Negative Binomial with mean $R_t$ and dispersion $k$. Smaller $k$ implies greater heterogeneity, with more infections concentrated in a few superspreading events.
- **Finite horizon**: Instead of assuming chains grow indefinitely, we truncate them by the number of generations that fit into the analysis window (e.g., ~18 generations over 45 days for influenza). This yields a finite expected chain size consistent with the time frame of interest.
- **Setting split**: We assume that a fraction of introductions occur in **micro-settings** (e.g., dense indoor gatherings) with higher $R_t$, while the remainder occur in the wider **community** where $R_t$ is lower. This reflects the heterogeneity of real-world transmission opportunities and highlights how a minority of introductions can drive most of the spread.

# Functions

```{r functions}
# -----------------------------
## 1) CDF of a Lognormal Incubation Period
##    To estimate the probability that an exposed traveler becomes
##    infectious while still in the destination during their stay.
##    What fraction of exposed travelers will have completed incubation (i.e., become infectious) within t days?
## -----------------------------
pincln <- function(t, meanlog = log(1.5), sdlog = 0.5) {
    # Arguments:
  #   t       : time (days) since exposure
  #   meanlog : mean of log(incubation)
  #   sdlog   : sd of log(incubation)
  # Returns:
  #   P(Incubation time <= t)
  # NOTE: I default to a median around exp(meanlog) ≈ 1.5 days,
  #       consistent with short influenza incubation.
  plnorm(t, meanlog = meanlog, sdlog = sdlog)
}

## -----------------------------
## 2) Expected Total Size Up To G Generations
##    Branching-process expectation over a finite horizon. If R<1,
##    total infinite expectation is 1/(1-R); but I cap at G generations
##    to reflect a limited analysis window (e.g., 45 days in summer).
## -----------------------------
expected_size_truncated <- function(R, G) {
   # Arguments:
  #   R : effective reproduction number for the setting
  #   G : integer number of generations included in horizon
  # Returns:
  #   E[total cases per seed] over generations 0..G
  if (abs(R - 1) < 1e-12) return(G + 1) # special case R==1
  (1 - R^(G + 1)) / (1 - R) # geometric sum
}

## -----------------------------
## 3) Extinction Probability for NB Offspring
##    When R>1, there’s a nonzero chance a seed dies out due to
##    stochasticity/overdispersion. I solve s = G(s) where G is the
##    PGF of the NegBin offspring distribution with mean=R, dispersion=k.
##    This yields the per-seed extinction probability q in [0,1].
## -----------------------------
extinction_prob_nb <- function(R, k, tol = 1e-10, maxit = 10000) {
  # Arguments:
  #   R   : mean offspring number
  #   k   : dispersion (smaller -> more overdispersion)
  #   tol : fixed-point tolerance
  #   maxit: max iterations for the fixed-point solver
  # Returns:
  #   q_ext : extinction probability. If R<=1, extinction is certain (=1).
  if (R <= 1) return(1.0)
    # PGF for NB(size=k, mean=R): G(s) = (1 + (R/k)*(1 - s))^(-k)
  s <- 0.99 # start near 1 for numerical stability
  for (i in 1:maxit) {
    Gs <- (1 + (R / k) * (1 - s))^(-k)
    if (abs(Gs - s) < tol) break
    s <- Gs
  }
  pmin(pmax(s, 0), 1)
}

## -----------------------------
## 4) CORE CALCULATOR
##    WHY: Glue together the ingredients—compute seeds, split them by
##    setting, compute expected totals (deterministic), and quantify risk
##    of ≥1 self-sustaining chain in micro-settings.
## -----------------------------
importation_calculator <- function(
#    travelers            = 300000,   # total arrivals during analysis window
    travelers            = 3000000,   # total arrivals during analysis window
    p_inf_on_arrival     = 0.01,     # point prevalence: infectious at arrival
    p_exposed_on_arrival = 0.005,    # arriving incubating (pre-symptomatic)
    stay_days            = 30,       # mean stay duration in destination
    inc_meanlog          = log(1.5), # incubation lognormal params
    inc_sdlog            = 0.5,
    gen_time_days        = 2.5,      # mean generation time (flu: ~2–3 d)
    horizon_days         = 45,       # observation window for onward spread
    frac_micro_setting   = 0.15,     # fraction of seeds entering high-R micro-settings
    R_comm               = 0.8,      # community Rt (often <1 in summer)
    R_micro              = 1.2,      # micro-setting Rt (can be >1 indoors)
    k_disp               = 0.3,      # NegBin dispersion (smaller -> more heterogeneity)
    simulate             = FALSE,    # toggle Monte Carlo
    nsims                = 2000,     # number of simulations (per setting)
    max_gen_cap          = 50        # safety cap on simulated generations
) {

  # ---- Probability that an EXPOSED traveler becomes infectious while still staying ----
  # CDF at 'stay_days' answers: "What fraction of the exposed will flip
  # to infectious before they leave?"
  #p_inc_during_stay <- plnorm(stay_days, meanlog = inc_meanlog, sdlog = inc_sdlog)
  p_inc_during_stay <- pincln(stay_days, meanlog = inc_meanlog, sdlog = inc_sdlog)  
  # ---- Expected number of infectious "seeds" contributed by travelers ----
  # Seeds come from two channels:
  #   (a) Infectious at arrival
  #   (b) Exposed at arrival who become infectious during their stay
  seeds_expected <- travelers * (p_inf_on_arrival + p_exposed_on_arrival * p_inc_during_stay)

  # ---- Split seeds between community vs. micro-settings ----
  # Rationale: some introductions will occur in settings where Rt>1
  # (e.g., crowded indoor workplaces, camps), even if the community-wide Rt<1.
  seeds_micro <- seeds_expected * frac_micro_setting
  seeds_comm  <- seeds_expected - seeds_micro

  # ---- Convert horizon (days) into generations ----
  # We cap the expected growth by the number of generations that fit in the
  # time horizon to avoid accumulating unrealistic long chains in summer.
  G <- floor(horizon_days / gen_time_days)

  # ---- Deterministic expected totals via truncated geometric sums ----
  E_size_comm  <- expected_size_truncated(R_comm,  G)
  E_size_micro <- expected_size_truncated(R_micro, G)

  expected_total_comm  <- seeds_comm  * E_size_comm
  expected_total_micro <- seeds_micro * E_size_micro
  expected_total_all   <- expected_total_comm + expected_total_micro

  # ---- Probability of ≥1 self-sustaining chain in micro-settings ----
  # For R_micro>1, some seeds ignite large chains; with NB offspring,
  # extinction probability per seed is q. With N independent seeds,
  # P(no self-sustaining chain) ≈ exp(-N*(1-q)) if seeds are rare/independent.
  q_ext_micro <- extinction_prob_nb(R_micro, k_disp)
  prob_ge1_self_sustaining_micro <- if (R_micro > 1) 1 - exp(-seeds_micro * (1 - q_ext_micro)) else 0

  # ---- Prepare output ----
  out <- list(
    seeds = list(
      expected_total_seeds = seeds_expected,
      seeds_comm = seeds_comm,
      seeds_micro = seeds_micro
    ),
    deterministic = list(
      generations = G,
      E_size_comm_per_seed  = E_size_comm,
      E_size_micro_per_seed = E_size_micro,
      expected_total_comm   = expected_total_comm,
      expected_total_micro  = expected_total_micro,
      expected_total_all    = expected_total_all
    ),
    risk = list(
      extinction_prob_micro = q_ext_micro,
      prob_ge1_self_sustaining_micro = prob_ge1_self_sustaining_micro
    )
  )

  # ---- Optional: Monte Carlo to capture variability in per-seed sizes ----
  # We simulate NB offspring branching processes for a single seed,
  # for G generations, and summarize the per-seed size distribution.
  if (simulate) {
    # Draw NB offspring counts with mean=R and dispersion=k
    r_offspring <- function(n, meanR, k) rnbinom(n, size = k, mu = meanR)

    # Simulate one seed’s total size over <= G generations
    simulate_branch <- function(R, k, G, cap = max_gen_cap) {
      total <- 1L # includes the index case
      gen <- 0L
      frontier <- 1L # active infectious nodes this generation
      while (gen < G && frontier > 0L && total < 1e7 && gen < cap) {
        new <- sum(r_offspring(frontier, R, k))
        total <- total + new
        frontier <- new
        gen <- gen + 1L
      }
      total
    }

    n_comm  <- max(0L, as.integer(round(seeds_comm)))
    n_micro <- max(0L, as.integer(round(seeds_micro)))

    comm_sizes  <- if (n_comm  > 0) replicate(nsims, simulate_branch(R_comm,  k_disp, G)) else numeric(0)
    micro_sizes <- if (n_micro > 0) replicate(nsims, simulate_branch(R_micro, k_disp, G)) else numeric(0)

    out$monte_carlo <- list(
      mean_size_per_seed_comm  = if (length(comm_sizes))  mean(comm_sizes)  else NA_real_,
      mean_size_per_seed_micro = if (length(micro_sizes)) mean(micro_sizes) else NA_real_,
      p95_size_per_seed_comm   = if (length(comm_sizes))  as.numeric(quantile(comm_sizes, 0.95))  else NA_real_,
      p95_size_per_seed_micro  = if (length(micro_sizes)) as.numeric(quantile(micro_sizes, 0.95)) else NA_real_
    )
  }

  class(out) <- "importation_calc"
  out
}

# -----------------------------
## 5) Pretty Printer
##    WHY: Summarize results in a readable block without needing to
##    remember element names in the output list.
## -----------------------------
print.importation_calc <- function(x, ...) {
  cat("\n================  Importation Calculator  ================\n")
  cat("Deterministic expectation over finite horizon\n")
  cat("----------------------------------------------------------\n")
  cat(sprintf("Expected infectious seeds (all): %.0f\n", x$seeds$expected_total_seeds))
  cat(sprintf("  • Community seeds: %.0f\n", x$seeds$seeds_comm))
  cat(sprintf("  • Micro-setting seeds: %.0f\n", x$seeds$seeds_micro))
  cat(sprintf("Generations within horizon: %d\n", x$deterministic$generations))
  cat(sprintf("Per-seed expected size (community): %.2f\n", x$deterministic$E_size_comm_per_seed))
  cat(sprintf("Per-seed expected size (micro)    : %.2f\n", x$deterministic$E_size_micro_per_seed))
  cat(sprintf("Expected cases (community): %.0f\n", x$deterministic$expected_total_comm))
  cat(sprintf("Expected cases (micro)    : %.0f\n", x$deterministic$expected_total_micro))
  cat(sprintf("Expected cases (TOTAL)    : %.0f\n", x$deterministic$expected_total_all))
  cat("\nSelf-sustaining chain risk in micro-settings (if R_micro>1)\n")
  cat("----------------------------------------------------------\n")
  cat(sprintf("Extinction prob per seed (micro): %.3f\n", x$risk$extinction_prob_micro))
  cat(sprintf("Prob(≥1 self-sustaining chain)  : %.3f\n", x$risk$prob_ge1_self_sustaining_micro))
  if (!is.null(x$monte_carlo)) {
    cat("\nMonte Carlo (per-seed size distribution)\n")
    cat("----------------------------------------------------------\n")
    cat(sprintf("Mean per-seed size (community):  %.2f | p95: %.0f\n",
                x$monte_carlo$mean_size_per_seed_comm, x$monte_carlo$p95_size_per_seed_comm))
    cat(sprintf("Mean per-seed size (micro)    :  %.2f | p95: %.0f\n",
                x$monte_carlo$mean_size_per_seed_micro, x$monte_carlo$p95_size_per_seed_micro))
  }
  cat("\n==========================================================\n")
  invisible(x)
}
```

# Baseline Scenario: 3 Million Visitors

```{r baseline, results='markup'}
## ============================
## EXAMPLE (toggle to run)
## ============================
 set.seed(42)
 res <- importation_calculator(
#   travelers            = 300000,   # arrivals from Southern Cone in June
   travelers            = 3000000,   # arrivals from Southern Cone in June
   p_inf_on_arrival     = 0.010,    # 1% infectious on arrival
   p_exposed_on_arrival = 0.005,    # 0.5% incubating
   stay_days            = 30,       # ~1 month
   gen_time_days        = 2.5,      # influenza generation time
   horizon_days         = 45,       # track for ~1.5 months
   frac_micro_setting   = 0.15,     # fraction in higher-risk indoor settings
   R_comm               = 0.80,     # summer community Rt (subcritical)
   R_micro              = 1.20,     # micro-environment Rt (>1)
   k_disp               = 0.30,     # overdispersion
   simulate             = TRUE,     # show variability
   nsims                = 2000
 )
 print(res)
 
```

# Sensitivity Analysis

```{r sensitivity}
# ============================
# Sensitivity Analysis
# ============================

# Create a grid of parameter combinations we want to test.
# - travelers: number of incoming visitors (1M, 3M, 6M)
# - R_micro:   effective reproduction number in micro-settings (1.1, 1.2, 1.4)
# - frac_micro_setting: fraction of introductions that occur in micro-settings
#   (10%, 15%, 20%)
# KEEP.OUT.ATTRS = FALSE removes extra attributes from expand.grid output
# stringsAsFactors = FALSE ensures numeric values stay numeric (not factors)
param_grid <- expand.grid(
  travelers = c(1e6, 3e6, 6e6),
  R_micro   = c(1.1, 1.2, 1.4),
  frac_micro_setting = c(0.10, 0.15, 0.20),
  KEEP.OUT.ATTRS = FALSE, stringsAsFactors = FALSE
)

# Loop over each row of the parameter grid
# seq_len(nrow(param_grid)) ensures we cover all rows safely
sweep_res <- lapply(seq_len(nrow(param_grid)), function(i) {
  
  # Extract the i-th parameter set (as a 1-row data.frame)
  p <- param_grid[i, ]
  
  # Run the importation calculator with that parameter set.
  # simulate = FALSE means we skip the Monte Carlo and just compute deterministic expectations
  out <- importation_calculator(
    travelers = as.numeric(p$travelers),           # number of travelers in this scenario
    R_micro   = as.numeric(p$R_micro),             # micro-setting reproduction number
    frac_micro_setting = as.numeric(p$frac_micro_setting), # fraction of seeds in micro-settings
    simulate  = FALSE
  )
  
  # Return a tidy 1-row data.frame summarizing results for this parameter set:
  # - travelers: number of incoming travelers
  # - R_micro:   micro-setting reproduction number
  # - frac_micro_setting: fraction of seeds in micro-settings
  # - seeds_all: expected number of infectious seeds generated
  # - cases_total: expected total cases over the horizon
  # - prob_self_sust_micro: probability of ≥1 self-sustaining chain in micro-settings
  data.frame(
    travelers = as.numeric(p$travelers),
    R_micro   = as.numeric(p$R_micro),
    frac_micro_set = as.numeric(p$frac_micro_setting),
    seeds_all = out$seeds$expected_total_seeds,
    cases_total = out$deterministic$expected_total_all,
    prob_self_sust_micro = out$risk$prob_ge1_self_sustaining_micro
  )
})

# Combine all the individual 1-row data.frames into a single table
sweep_tbl <- do.call(rbind, sweep_res)

# Print the sensitivity table
sweep_tbl
```
